---
title: "Feature Selection and Prediction of Substance Use and Reckless Behaviors in YRBS 2019 Data using Lasso"
author: "Catalina Canizares"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Trees}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
suppressPackageStartupMessages(library(tidymodels))
library(MLearnYRBSS)
```

Welcome to the first vignette of our educational data package `MLearnYRBSS` on supervised machine learning methods with `tidymodels`. In this vignette, we aim to investigate the research question of whether we can predict marijuana use based on various predictors. These predictors include alcohol and other drug use, reckless behaviors such as not wearing seatbelts or driving under the influence, as well as variables related to sexual abuse, carrying guns to school, bullying, and feeling unsafe while walking to school. 

# Explore the Data
```{r}

```

